---
title: "Data Mining - project (Adults)"
author: "Wiktoria Fimińska 262283, Julia Grzegorzewska 262314"
date: "`r Sys.Date()`"
fontsize: 12pt
output: 
  pdf_document:
        toc: true
        toc_depth: 3
        number_sections: true
        fig_caption: true
        
header-includes:
    - \usepackage{polski}
    - \usepackage[utf8]{inputenc}
    - \usepackage{graphicx}
    -  \usepackage{amsmath}
    - \usepackage{multirow}
    - \usepackage{float}
    - \usepackage{enumitem}
    - \usepackage{mathtools}
    - \usepackage{hyperref}
    - \usepackage{url}
    - \mathtoolsset{showonlyrefs}
    - \usepackage{caption}
    - \usepackage{geometry}
    
---
\newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

```{r echo=FALSE}
pdf.options(encoding = 'CP1250')
```

\section{Introduction}


\section{Data}
```{r, echo = FALSE, message=FALSE}
library(dplyr)
library(DataExplorer)
data <- read.csv(file="data\\adult.data", stringsAsFactors = TRUE, col.names=c("age", "workclass", "fnlwgt", "education", 
                                                       "education.num","marital.status", "occupation", "relationship", "race","sex",
                                                       "capital.gain", "capital.loss", "hours.per.week","native.country", "income"))
attach(data)
data %>% sample_n(5)
```

```{r}
# which columns have '?' values
columns_with_question <- names(data)[sapply(data, function(col) any(col == " ?"))]

print(columns_with_question)
```


\subsection{Data characteristics}
Dataset contains $14$ variables, which are:
\begin{enumerate}[label=(\alph*)]
    \item \textbf{age}\\
    age of a person - continuous variable.
    \item \textbf{workclass}\\
    workclass of a person: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked.
    \item \textbf{fnlwgt}\\
    NO IDEA WHAT IT IS
    \item \textbf{education}\\
    level of education of a person: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.
    \item \textbf{education-num}\\
    continuous version of education (ONE OF THEM WON'T BE NECESSARY TO USE - WE JUST DROP IT)
    \item \textbf{marital-status}\\
    marital status of a person: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
    \item \textbf{occupation}\\ 
    occupation of a person: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces.
    \item \textbf{relationship}\\
    relationship status of a person: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried.
    \item \textbf{race}\\
    race of a person: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black.
    \item \textbf{sex}\\
    sex of a person: Female, Male.
    \item \textbf{capital-gain}\\
    amount of money a person gained - continuous variable.
    \item \textbf{capital-loss}\\
    amount of money a person loosed - continuous variable.
    \item \textbf{hours-per-week}\\
    XYZ idk
    \item \textbf{native-country}\\
    native-country of a person: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad\&Tobago, Peru, Hong, Holand-Netherlands.
\end{enumerate}

\subsection{Data preparation}
```{r, echo = FALSE}
columns_with_question <- sapply(data, function(col) sum(col == ' ?') > 0)
for (column in names(data)[columns_with_question]) {
  data[data[[column]] == ' ?', column] <- NA
}
data %>% sample_n(10)
```

Missing observations?
```{r}
sum(is.na(data))
```
WE NEED TO FIND A METHOD TO FILL THE NANS 

```{r}
summary(age)
summary(capital.gain)
summary(capital.loss)
summary(hours.per.week)
```

```{r}
table(workclass)
table(education)
table(marital.status)
table(occupation)
table(relationship)
table(race)
table(sex)
table(income)
table(native.country)
```
```{r, fig.cap="Charts: continuous quantitative data", fig.align="center"}
par(mfrow=c(2,3))
hist(age)
hist(hours.per.week)
hist(capital.gain)
boxplot(age, main="Age")
boxplot(hours.per.week, main="Hours per week")
boxplot(capital.gain, main="Capital loss")
par(mfrow=c(1,1))
```

\section{Analysis}


\section{Discussion}
